Procurements of the Canadian Government, Runaway Military Contracts

### 2024-12-03 paper/paper.qmd
fixed top suppliers graph

df <- read_csv(here::here("data/analysis_data/cleaned_data.csv"))
aggregated_df <- read_csv(here::here("data/analysis_data/top_ten_suppliers.csv"))
model <- read_rds(here::here("models/gen_model.rds"))
mil_model <- read_rds(here::here("models/mil_model.rds"))
```

@@

colors <- c("#377eb8", "#e41a1c", "#4daf4a", "#ff7f00", "#984ea3", "#808080")

@@

# TODO: Assign specific colors to each buyer group
# Example color assignments (replace with desired colors)
buyer_colors <- c(
  "National Defence" = "#e41a1c",
  "Public Services and Procurement Canada" = "#984ea3",
  "Shared Services Canada" = "#ff7f00",
  "Public Health Agency of Canada" = "#4daf4a",
  "Fisheries and Oceans Canada" = "#377eb8",
  "Other"  = "#808080"
)

@@

# Create the bar chart
ggplot(data = military_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "#e41a1c") +
  labs(x = "Year", y = "Total Spending (in Billions)") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

# Create the bar chart
ggplot(data = public_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "#984ea3") +
  labs(x = "Year", y = "Total Spending (in Billions)") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r fig-top-buyers}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center
#| fig-pos: "H"

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending_billion, fill = buyer_category)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.9), # Adjust dodge width
    width = 0.8 # Adjust bar width
  ) +
  scale_fill_manual(
    values = c(
      "National Defence" = "steelblue",
      "Public Services and Procurement Canada" = "orange",
      "Shared Services Canada" = "green",
      "Public Health Agency of Canada" = "purple",
      "Fisheries and Oceans Canada" = "red",
      "Other" = "gray"
    )
  ) +
  scale_y_continuous(
    labels = comma_format(accuracy = 1), # Formats y-axis labels without decimal places
    breaks = pretty_breaks(n = 5), # Ensures approximately 5 breaks like 5, 10, 15
    expand = expansion(mult = c(0, 0.05)) # Adds a small space above the highest bar
  ) +
  labs(
    title = "Spending by Selected and Other Buyers Per Year",
    subtitle = "Amounts are in Billions of CAD",
    x = "Year",
    y = "Total Spending (Billions)",
    fill = "Buyer",
    caption = "."
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for readability
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.caption = element_text(hjust = 0.5, size = 10, face = "italic") # Centered caption
  )

@@

```{r fig-top-supps}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center
#| fig-pos: "H"

#TODO Fix colors and supplier names and order of supps (highest to lowest)

# Create the bar chart with legend removed
ggplot(aggregated_df, aes(x = reorder(word, total_amount), y = total_amount, fill = buyer)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Top 10 Suppliers by Aggregated Amount Awarded",
    x = "Supplier",
    y = "Total Amount Awarded"
    # Remove the legend title by not specifying 'fill' in labs
  ) +
  theme_minimal() +
  theme(legend.position = "none")

>>>>

df <- read_csv(here::here("data/analysis_data/cleaned_data.csv"))
aggregated_df <- read_csv(here::here("data/analysis_data/top_ten_suppliers.csv"))
model <- read_rds(here::here("models/gen_model.rds"))
mil_model <- read_rds(here::here("models/mil_model.rds"))

#colors!
def_c <- "#e41a1c"
fish_c <- "#377eb8"
health_c <- "#4daf4a"
services_c <- "#ff7f00"
public_c <- "#984ea3"
space_c <- "#ffff33"

@@

@@

# TODO: Assign specific colors to each buyer group
# Example color assignments (replace with desired colors)
buyer_colors <- c(
  "National Defence" = def_c,
  "Public Services and Procurement Canada" = public_c,
  "Shared Services Canada" = services_c,
  "Public Health Agency of Canada" = health_c,
  "Fisheries and Oceans Canada" = fish_c,
  "Other"  = "gray"
)

@@

# Create the bar chart
ggplot(data = military_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = def_c) +
  labs(x = "Year", y = "Total Spending (in Billions)") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

# Create the bar chart
ggplot(data = public_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = public_c) +
  labs(x = "Year", y = "Total Spending (in Billions)") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r fig-top-buyers}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Spending by Top 5 Buyers and Others Per Year
#| fig-align: center
#| fig-pos: "H"

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending_billion, fill = buyer_category)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.9), # Adjust dodge width
    width = 0.8 # Adjust bar width
  ) +
  scale_fill_manual(
    values = c(
  "National Defence" = def_c,
  "Public Services and Procurement Canada" = public_c,
  "Shared Services Canada" = services_c,
  "Public Health Agency of Canada" = health_c,
  "Fisheries and Oceans Canada" = fish_c,
  "Other"  = "gray"
    )
  ) +
  scale_y_continuous(
    labels = comma_format(accuracy = 1), # Formats y-axis labels without decimal places
    breaks = pretty_breaks(n = 5), # Ensures approximately 5 breaks like 5, 10, 15
    expand = expansion(mult = c(0, 0.05)) # Adds a small space above the highest bar
  ) +
  labs(
    x = "Year",
    y = "Total Spending (Billions)",
    fill = "Buyer",
    caption = "."
  ) +
  theme_minimal() +
  theme(
    legend.text = element_text(size = 8), # Adjusts the size of legend text
    legend.title = element_text(size = 9), # Adjusts the size of legend title
    legend.key.size = unit(0.6, "lines") # Reduces the size of legend symbols
  )

@@

```{r fig-top-supps}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Top 10 Suppliers by Total Amount Awarded
#| fig-align: center
#| fig-pos: "H"
# Merge total aggregated amounts back into aggregated_df
aggregated_df <- aggregated_df %>%
  left_join(aggregated_totals, by = "word", suffix = c("", "_total"))

# Get the top 10 suppliers
top_suppliers <- aggregated_totals %>%
  top_n(10, total_amount)

# Filter aggregated_df to include only the top 10 suppliers
aggregated_df_top10 <- aggregated_df %>%
  filter(word %in% top_suppliers$word)

# Use this ordering in the plot
# Use this ordering in the plot (reversed order)
ggplot(aggregated_df_top10, aes(x = reorder(word, total_amount_total), y = total_amount / 1e9, fill = buyer)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    x = "Supplier",
    y = "Total Amount Awarded (in Billions)",
    fill = "Buyer"  # Adds a title to the legend
  ) +
  scale_y_continuous(labels = scales::comma) +  # Format y-axis with commas
  scale_fill_manual(
    values = c(
      "National Defence" = def_c, 
      "Canadian Space Agency" = space_c, 
      "Shared Services Canada" = services_c, 
      "Public Health Agency of Canada" = health_c, 
      "Fisheries and Oceans Canada" = fish_c,
      "Other" = "gray"  # Assign a default color for other buyers if necessary
        ), 
    breaks = selected_buyers  # Restricts legend to selected buyers
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"  # Positions the legend on the right
  )
